---
title: "Statistical assignment 1"
author: "Eliot Barrett - Holman 68008915"
date: "29/01/20"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Open data (10 points)

In this assignment you will work with the individual level data from wave 8 of the Understanding Society survey. First, you need to open the data set. Please complete the code below.

```{r}
library(tidyverse)
Data <- read_tsv("C:/Users/eliot/Documents/Politics and IR/Year 2 Term 2/Data 3/Data3_blank/data/UKDA-6614-tab/tab/ukhls_w8/h_indresp.tab")
```

Now you have got your data frame stored as Data. 

## Select variables (10 points)

The data for Wave 8 of the Understanding Society were collected in 2016-18. Among other things, people were asked the following question: "Should the United Kingdom remain a member of the European Union or leave the European Union?" In this assignment, we will explore how answers to this question depend on sex and age.

First, you need to select the variables for the analysis. You want to keep the following variables: cross-wave individual identifier (*pidp*), support for the UK remaining or leaving the EU (*h_eumem*), sex (*h_sex_dv*), age (*h_age_dv*), and sample origin (*h_memorig*).

Complete the code below to select those variables from the data frame and save the result.

```{r}
Data <- Data %>%
        select(pidp, h_eumem, h_sex_dv, h_age_dv, h_memorig)
```

## Filter observations (10 points)

To make nationally representative estimates from the Understanding Society data we would need to use weight coefficients. There are many different types of weight coefficients that can be used depending on the question and the level of analysis (see the User Guide, pp. 65-71). We will not do this in this assignment. However, what we want to do is to keep data from the original Understanding Society sample only (ukhls gb 2009-10), dropping  data for Northern Ireland, the BHPS cohort members and ethnic minority boost samples. This will make data closer to be representative for Great Britain. You need to choose the observations where *h_memorig* has the value of 1.

```{r}
Data <- Data %>%
        filter(h_memorig == 1)

table(Data$h_memorig)
```

## Recode data (20 points)

Let us tabulate the variables for EU support, sex, and age.

```{r}
table(Data$h_eumem)
table(Data$h_sex_dv)
table(Data$h_age_dv)
```

You will see that all these variables are numeric. You can learn what the numeric codes mean by checking the codebook here: https://www.understandingsociety.ac.uk/documentation/mainstage/dataset-documentation/datafile/indresp/wave/8 .

We want to do the following:

1) Recode the variable for EU support as binary (1 for Remain, 0 for Leave), coding all types of missing values (including refusals and "don't know") as NA.
2) Recode sex into a character vector with the values "male" or "female".
3) Recode age into a variable with the following categories: 16 to 25, 26 to 40, 41 to 55, 56 to 70, over 70.

In all cases, we want to create new variables.

```{r}
Data <- Data %>%
        mutate(EU = recode(h_eumem, 
                           `1` = 1,
                           `2` = 0,
                           .default = NA_real_)
      
        ) %>%
        mutate(sex = recode(h_sex_dv,
                            `1` = "male",
                            `2` = "female",
                            .default = NA_character_)
                            
                            
                            )  %>%
        mutate(agegr = case_when(
                                  between(h_age_dv, 16, 25) ~ "16 to 25",
                                  between(h_age_dv, 26, 40) ~ "26 to 40",
                                  between(h_age_dv, 41, 55) ~ "41 to 55",
                                  between(h_age_dv, 56, 70) ~ "56 to 70",
                                  h_age_dv > 70 ~ "Over 70"
        )
        )




```

## Summarise data (20 points)

Let us **dplyr** to calculate how many people in the sample supported Remain and Leave, both as absolute numbers and percentages.

```{r}
        Data %>%
  count(EU) %>%
  mutate(perc = n / sum(n) * 100)
        
```

Write a couple of sentences with the interpretation of this result. How this compares with the result of the 2016 referendum? Why?
This result shows a 48% remain vote compared to a 41% leave vote, this is opposite to the results of the 2016 referendum which saw a leave majority.
Potential explanations for this difference can be either social desirability bias, methodology of the sample or even that a significant proportion who were sampled did not use their vote
Moreover, it is likely that there are a large number of those who voted leave in the referendum who have been caught in the missing data

## Summarise data by sex and age (30 points)

Now let us look at the support for Leave and Remain by sex and age. Use your newly created variables.


```{r}

Data %>%  group_by(sex, agegr) %>%
  summarise(
    percvotes = mean(EU, na.rm = TRUE) * 100
  )
        
        
```


Write a couple of sentences interpreting your results.
Here we see a breakdown of proportion voting remain for each group, split by age group and sex.
We see females aged 16-25 being most likely to vote remain (73%), followed by the male 16-25 group (66%)
On the other side of the spectrum we see that the male over 70 group is least likely to support remain (38%), followed by the over 70 female group (44%).
Indeed, this confirms that young people were more likely to support remain while the older generations were less likely to do so.
Interestingly, as a whole we see females as more likely to support remain than their male counterparts across each age group